Conversations in time: interactive visualisation to explore structured temporal data

### 2020-09-23 tsibbletalk.Rmd
done with linking section

The tree plot provides the data skeleton, upon which the rest of data can be attached. In this example, small multiples of line plots are placed at the top right of Figure \ref{fig:tourism-linking-fig} to explore the temporal trend across regions, by trip purpose. The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code as follows:

@@

To tease apart these overlaid time series, they are funnelled through the `features()` S3 method to extract some key characteristics, including the measurements of trend and seasonality. A scatterplot is populated from these statistics for each series.

@@

Lastly, three graphics are composed as an ensemble of coordinated views for multi-facetted exploration, shown as Figure \ref{fig:tourism-linking-fig} (the interactive realisation of Figure \ref{fig:highlight-retail}). Routine functions bring about new interaction with temporal data on the client side.

@@

Since all plots are stemmed from one shared tsibble data source, they are self-linking views. Nodes, lines, and points are hoverable and clickable. Given the `spec`, clicking either one element in any plot highlights all points that match the `Region` category, briefly "categorical linking". In Figure \ref{fig:tourism-linking-fig}, when hovering and selecting the circle associated with "Sydney" in the scatter plot, all data records with shared values of "Sydney" listen and react to this interaction via self updating in red. In order for comparison with other regions or states, press the "Shift" key to enable persistent selection, and simultaneously select the parent node on the tree, saying "Western Australia", to include all the children by switching to the blue colour. The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over years. Despite of relatively weaker performance in Western Australia, Australia's North West region sees the strongest upward trend, bypassing Sydney in some years.

In summary, shared tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data over their common term "key". The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linkings. And the `plotly_key_tree()` in turn decodes the specification to plot a tree for data overview and navigation, accompanied with more detailed plots.

>>>>

The tree plot provides the graphics skeleton, upon which the rest of data plots can be attached. In this example, small multiples of line plots are placed at the top right of Figure \ref{fig:tourism-linking-fig} to explore the temporal trend across regions, by trip purpose. The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code to create this.

@@

These line plots are heavily overplotted. To tease apart structure in the multiple time series, the `features()` function `fabletools` computes interesting characteristics, including the measures of trend and seasonality. These are displayed in the scatterplot at bottom right, where one dot represents one series.

@@

There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \ref{fig:tourism-linking-fig}.  (This is the interactive realisation of Figure \ref{fig:highlight-retail}). <!-- Routine functions bring about new interaction with temporal data on the client side.-->

@@

Since all plots are created from one shared tsibble data source, they are self-linking views. Nodes, lines, and points are hoverable and clickable. Given the `spec`, clicking either one element in any plot highlights all points that match the `Region` category, that is, categorical linking. Figure \ref{fig:tourism-linking-fig} is a static view of an interactive exploration. The steps in getting to this point were:

1. A branch of the tree corresponding to Western Australia was first selected. (The names of the regions are a little odd, which is a quirk of the data set, but all four areas, Australia's South West, ...., correspond to tourist destinations in Western Australia. Hovering over the node on the branch brings up the state name.) This generated the response in the line plots and the scatterplot that colored corresponding time series and points as blue. 
2. To enable persistent selection, in oder to compare regions or states, "Shift" and click on the tree was done, after switching the colour to red. This generated the response that points and time series corresponding to Sydney were highlighted in red. 
3. Hovering over the points brings up the label for Sydney.

<!-- In Figure \ref{fig:tourism-linking-fig}, when hovering and selecting the circle associated with "Sydney" in the scatter plot, all data records with shared values of "Sydney" listen and react to this interaction via self updating in red. In order for comparison with other regions or states, press the "Shift" key to enable persistent selection, and simultaneously select the parent node on the tree, saying "Western Australia", to include all the children by switching to the blue colour.-->
The domestic tourism sees Sydney as one of the most popular destinations in realm of business and friends visiting over the years. Despite the relatively weaker performance in Western Australia, Australia's North West region sees a strongest upward trend in Business, bypassing Sydney in some years.

In summary, shared the tsibble data nicely bridges between the \CRANpkg{crosstalk} and **tidyverts** ecosystems for temporal data using the common "key". The `as_shared_tsibble()` provides a symbolic user interface for effortless construction of a hybrid of hierarchical and categorical linking between plots. The `plotly_key_tree()` fuction, in turn, decodes the hierarchical specification to plot a tree for data overview and navigation, when accompanied by more detailed plots.

### 2020-09-23 tsibbletalk.tex
done with linking section

The tree plot provides the data skeleton, upon which the rest of data
can be attached. In this example, small multiples of line plots are
placed at the top right of Figure \ref{fig:tourism-linking-fig} to
explore the temporal trend across regions, by trip purpose. The shared
tsibble data can be directly piped into \CRANpkg{ggplot2} code as
follows.

@@

To tease apart these overlaid time series, they are funnelled through
the \texttt{features()} S3 method to extract some key characteristics,
including the measurements of trend and seasonality. A scatterplot is
populated from these statistics for each series.

@@

Lastly, three graphics are composed as an ensemble of coordinated views
for multi-facetted exploration, shown as Figure
\ref{fig:tourism-linking-fig} (the interactive realisation of Figure
\ref{fig:highlight-retail}). Routine functions bring about new
interaction with temporal data on the client side.

@@

Since all plots are stemmed from one shared tsibble data source, they
are self-linking views. Nodes, lines, and points are hoverable and
clickable. Given the \texttt{spec}, clicking either one element in any
plot highlights all points that match the \texttt{Region} category,
briefly ``categorical linking''. In Figure
\ref{fig:tourism-linking-fig}, when hovering and selecting the circle
associated with ``Sydney'' in the scatter plot, all data records with
shared values of ``Sydney'' listen and react to this interaction via
self updating in red. In order for comparison with other regions or
states, press the ``Shift'' key to enable persistent selection, and
simultaneously select the parent node on the tree, saying ``Western
Australia'', to include all the children by switching to the blue
colour. The domestic tourism sees Sydney as one of the most popular
destinations in realm of business and friends visiting over years.
Despite of relatively weaker performance in Western Australia,
Australia's North West region sees the strongest upward trend, bypassing
Sydney in some years.

In summary, shared tsibble data nicely bridges between the
\CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems for temporal data
over their common term ``key''. The \texttt{as\_shared\_tsibble()}
provides a symbolic user interface for effortless construction of a
hybrid of hierarchical and categorical linkings. And the
\texttt{plotly\_key\_tree()} in turn decodes the specification to plot a
tree for data overview and navigation, accompanied with more detailed
plots.

>>>>

The tree plot provides the graphics skeleton, upon which the rest of
data plots can be attached. In this example, small multiples of line
plots are placed at the top right of Figure
\ref{fig:tourism-linking-fig} to explore the temporal trend across
regions, by trip purpose. The shared tsibble data can be directly piped
into \CRANpkg{ggplot2} code to create this.

@@

These line plots are heavily overplotted. To tease apart structure in
the multiple time series, the \texttt{features()} function
\texttt{fabletools} computes interesting characteristics, including the
measures of trend and seasonality. These are displayed in the
scatterplot at bottom right, where one dot represents one series.

@@

There is one final step, to compose the three plots into an ensemble of
coordinated views for exploration, shown in Figure
\ref{fig:tourism-linking-fig}. (This is the interactive realisation of
Figure \ref{fig:highlight-retail}).

@@

Since all plots are created from one shared tsibble data source, they
are self-linking views. Nodes, lines, and points are hoverable and
clickable. Given the \texttt{spec}, clicking either one element in any
plot highlights all points that match the \texttt{Region} category, that
is, categorical linking. Figure \ref{fig:tourism-linking-fig} is a
static view of an interactive exploration. The steps in getting to this
point were:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A branch of the tree corresponding to Western Australia was first
  selected. (The names of the regions are a little odd, which is a quirk
  of the data set, but all four areas, Australia's South West, \ldots.,
  correspond to tourist destinations in Western Australia. Hovering over
  the node on the branch brings up the state name.) This generated the
  response in the line plots and the scatterplot that colored
  corresponding time series and points as blue.
\item
  To enable persistent selection, in oder to compare regions or states,
  ``Shift'' and click on the tree was done, after switching the colour
  to red. This generated the response that points and time series
  corresponding to Sydney were highlighted in red.
\item
  Hovering over the points brings up the label for Sydney.
\end{enumerate}

The domestic tourism sees Sydney as one of the most popular destinations
in realm of business and friends visiting over the years. Despite the
relatively weaker performance in Western Australia, Australia's North
West region sees a strongest upward trend in Business, bypassing Sydney
in some years.

In summary, shared the tsibble data nicely bridges between the
\CRANpkg{crosstalk} and \textbf{tidyverts} ecosystems for temporal data
using the common ``key''. The \texttt{as\_shared\_tsibble()} provides a
symbolic user interface for effortless construction of a hybrid of
hierarchical and categorical linking between plots. The
\texttt{plotly\_key\_tree()} fuction, in turn, decodes the hierarchical
specification to plot a tree for data overview and navigation, when
accompanied by more detailed plots.
